Using structural and evolutionary information to detect and correct pyrosequencing errors in non-coding RNAs

### 2013-07-01 JCB/results_JCB.tex
update text commnet ROC

{\centering \url{http://csb.cs.mcgill.ca/RNApyro}\\}
The time benchmarks were performed on an AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
Since typical use-cases of \RNApyro require efficiency and scalability, we present in TableFigure~\ref{fig:time}
typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
For those tests, both the multiple sequence alignment and the reference sequence were generated uniformly at random,
based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
\TODOVlad{Add more extensive time benchmark, as a plot}
%\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}

@@

\subsection{Correcting Illumina sequencing errors in 16s rRNAs}
\label{sec:16S}
\TODOTous{Describe and present results????}

@@

As in Section~\ref{sec:5S}, we evaluated the predictive power of \RNApyro-computed mutational profiles using a leave-one-out strategy. We picked a sequence at random from the MSA, sequenced/mutated it as above, and run \RNApyro to establish its mutational profile. In this execution, we used values of $\beta=15$ and $\alpha\in [0, 0.5, 0.8, 1.0]$, and set the presumed number of mutations to twice the average error rate made by {\tt ART}, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
We repeated the whole procedure $12$/$10$/$14$ times for the $5$/$10$/$15-$fold coverage. 
We evaluated the predictive power of the profile by a computing a joint ROC curve for each value of $\alpha$, and each coverage, as described in Section~\ref{sec:5S}. Figure~\ref{fig:16s_all} shows the ROC curves, computed over all positions, while Figure~\ref{fig:16s_bp} only focuses on positions that are paired in the consensus secondary structure.

 \begin{figure}
\centering
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/5fold-a}
  \caption{$5-$fold coverage}
\end{subfigure}%
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/10fold-a}
    \caption{$10-$fold coverage}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/15fold-a}
    \caption{$15-$fold coverage}
\end{subfigure}
\caption{Performance of error-correction over all positions. Subfigures show accuracy when
ART fold parameter is set to 5, 10 and $15-$fold coverage. We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values ranging of $\alpha = \{0, 0.5, 0.8, 1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over at least 10 experiments.}

@@

Our data demonstrates that even on long sequences our algorithm shows interesting potential for 
error-correction applications. First, the AUC values (up to 0.81) when looking at all positions, 
in Fig.~\ref{fig:16s_all},  indicate that a signal has been successfully extracted. When restraining
to structured regions (i.e. base paired regions) we obtain AUC values up to 0.998.  Since 
contributions to the energy and isostericity only arise from structured regions, this was an expected result.

An interesting feature is that best results are almost always when $\alpha=0$, i.e. all the contribution
comes from the isostericity. A notable exception is when the number of mutation is underestimated, 
in that case best performances were observed for non-null values of $\alpha$.

>>>>

{\centering \url{http://csb.cs.mcgill.ca/RNApyro}\\}
The time benchmarks were performed on an AMD Opteron(tm) Processor 6278  at 2.4 GHz with cache of 512 KB.20 cores and 175 GB ram.
Since typical use-cases of \RNApyro require efficiency and scalability, we present in TableFigure~\ref{fig:time}
typical runtimes required to compute the probabilities for  every nucleotide at every positions for a vast set of parameters.
For those tests, both the multiple sequence alignment and the reference sequence were generated uniformly at random,
based on a realistic random secondary structure, generated as described in Levin~\emph{et al}~\cite{Levin:2012kx}.
%\TODOVlad{Add more extensive time benchmark, as a plot}
%\TODOVlad{Is time benchmark still on Mac mini or mcbbigram?}

@@

\subsection{Correcting Illumina sequencing errors in 16s rRNAs}
\label{sec:16S}
%\TODOTous{Describe and present results????}

@@

As in Section~\ref{sec:5S}, we evaluated the predictive power of \RNApyro-computed mutational profiles using a leave-one-out strategy. We picked a sequence at random from the MSA, sequenced/mutated it as above, and run \RNApyro to establish its mutational profile. In this execution, we used values of $\beta=15$ and $\alpha\in [0, 0.5, 0.8, 1.0]$, and set the presumed number of mutations to twice the average error rate made by {\tt ART}, i.e. $4.8\%$, $1.8\%$ and $1.2\%$ for fold coverages of $5$, $10$ and $15$ respectively.
We repeated the whole procedure $30$ times for each ($5$/$10$/$15-$fold) coverage. 
We evaluated the predictive power of the profile by a computing a joint ROC curve for each value of $\alpha$, and each coverage, as described in Section~\ref{sec:5S}. Figure~\ref{fig:16s_all} shows the ROC curves, computed over all positions, while Figure~\ref{fig:16s_bp} only focuses on positions that are paired in the consensus secondary structure.

 \begin{figure}
\centering
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/5fold-a}
  \caption{$5-$fold coverage}
\end{subfigure}%
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/10fold-a}
    \caption{$10-$fold coverage}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
  \centering
  \includegraphics[width=\linewidth]{figures/15fold-a}
    \caption{$15-$fold coverage}
\end{subfigure}
\caption{Performance of error-correction over all positions. Subfigures show accuracy when
ART fold parameter is set to 5, 10 and $15-$fold coverage. We also analyze the impact of the parameter $\alpha$ distributing the weights of stacking pair energies vs isostericity scores and use values ranging of $\alpha = \{0, 0.5, 0.8, 1.0\}$. The AUC is indicated in the legend of the figures. Each individual ROC curve represent the average performance over 30 experiments.}

@@

Our data demonstrates that even on long sequences our algorithm shows interesting potential for 
error-correction applications. First, the AUC values (up to 0.81) when looking at all positions, 
in Fig.~\ref{fig:16s_all},  indicate that a signal has been successfully extracted. When restraining
to structured regions (i.e. base paired regions) we obtain AUC values up to 0.988.  Since 
contributions to the energy and isostericity only arise from structured regions, this was an expected result.

An interesting feature is that best results are almost always when $\alpha=0$, i.e. all the contribution
comes from the isostericity.
% A notable exception is when the number of mutation is underestimated, 
%in that case best performances were observed for non-null values of $\alpha$.
